---
title: "Introduction and exploring raw data"
author: "Andrew Hicks"
date: "`r Sys.Date()`"
output: html_document
---

```{r knitr-options, include = FALSE}
knitr::opts_chunk$set(fig.align="center")
```

setwd("C:/Users/Andrew Hicks/Documents/Analytics/GitHub/ContinuedLearning/DataCamp/R/Importing and Cleaning Data with R/Cleaning Data in R/")
weather <- readRDS("dataset/weather.rds")
bmi <- read.csv("dataset/bmi_clean.csv")
census_retail <- read.csv("dataset/census-retail.csv")
students_with_dates <- read.csv("dataset/students_with_dates.csv")

#### Data Science Process: Collect >> Clean >> Analyze >> Report
---
## What we'll cover in this course:
1. Exploring raw data
2. Tidying data
3. Preparing data for analysis
4. Putting it all together

---   

```{r}
# View the first 6 rows of data
head(weather)

# View the last 6 rows of data
tail(weather)

# View a condensed summary of the data
str(weather)
```

---

## Exploring RAW data:
1. **Understanding the structure of your data**

    ```
    # Load the lunch data

    lunch <- read.csv("dataset/lunch_clean.csv")
    ```

- **class() - Class of data object**

    ```
    class(lunch)

    > "data.frame"
    ```

- **dim() - Dimension of data**

    ```
    dim(lunch)

    > 46 7
    ```

- **names() - Column names**

    ```
    names(lunch)

    > "year" "avg_free" ... "perc_free_red"
    ```

- **str() - Preview of data with helpful details**

    ```
    str(lunch) # provides dataframe row & columns, column's names, datatype and heading records
    ```

- **glimpse() - Better version of str() from dplyr**

    ```
    # Load dplyr
    library(dplyr)
    glimpse(lunch)
    ```

- **summery() - Summary of data**
    
    ```  
    summary(lunch)
    ```

---

**2. Looking at your data**

- **head() - View top of dataset**

    ```
    head(lunch)
    ```
- **tail() - View bottom of dataset**

    ```
    tail(lunch)
    ```

- **print() - View entire dataset (not recommended!)**
  
- **hist() - View histogram of a single variable**

    ```
    hist(lunch$perc_free_red)
    ```
- **plot() - View plot of two variables**

    ```
    plot(lunch$year, lunch$perc_free_res)
    ```
---

**Practice**

    ```{r}
    # Print bmi to the console
    print(bmi)

    # View the first 6 rows
    head(bmi,6)

    # View the first 15 rows
    head(bmi,15)

    # View the last 6 rows
    tail(bmi,6)

    # View the last 10 rows
    tail(bmi,10)

    ```
    
---

**3. Visualize your data**

    ```

    # Histogram of BMIs from 2008
    hist(bmi$Y2008)

    # Scatter plot comparing BMIs from 1980 to those from 2008
    plot(bmi$Y1980, bmi$Y2008)

    ```

---

 

 
