---
title: "Preparing data for analysis"
author: "Andrew Hicks"
date: "`r Sys.Date()`"
output: html_document
---

```{r knitr-options, include = FALSE}
knitr::opts_chunk$set(fig.align="center")
```
setwd("C:/Users/Andrew Hicks/Documents/Analytics/GitHub/ContinuedLearning/DataCamp/R/Importing and Cleaning Data with R/Cleaning Data in R/")
weather <- readRDS("dataset/weather.rds")
bmi <- read.csv("dataset/bmi_clean.csv")
census-retail <- read.csv("dataset/census-retail.csv")
students_with_dates <- read.csv("dataset/students_with_dates.csv")

---

```{r}
class("hello")

class(factor("yes"))

class(3.844)

class(TRUE)

class(77L)
```

Type coversions:
```{r}
as.character(2016)

as.numeric(TRUE)

as.integer(99)

as.factor("something")
```

Lubridate Package: Coerce strings to dataset
```{r}
# Load the lubridate package
library(lubridate)

# Experiment with basic lubridate functions
ymd("2015-08-25")

ymd("2015 August 25")

mdy("August 25, 2015")

hms("13:33:09")

ymd_hms("2015/08/25 13:33:09")
```